---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%",
  dpi = 300,
  fig.align = "center"
)
#knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

# epinionDSB <img src='man/figures/logo.svg' align="right" height="139" />

<!-- badges: start -->
[![CodeFactor](https://www.codefactor.io/repository/github/jvieroe/epiniondsb/badge)](https://www.codefactor.io/repository/github/jvieroe/epiniondsb)
[![Github All Releases](https://img.shields.io/github/downloads/jvieroe/epinionDSB/total.svg)]()
[![](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License: GPL (>= 3)](https://img.shields.io/badge/license-GPL (>= 3)-blue.svg)](https://cran.r-project.org/web/licenses/GPL (>= 3))
[![](https://img.shields.io/github/last-commit/jvieroe/epinionDSB.svg)](https://github.com/jvieroe/epinionDSB/commits/main)
[![Travis build status](https://travis-ci.org/jvieroe/epinionDSB.svg?branch=main)](https://travis-ci.org/jvieroe/epinionDSB)

Master branch status: [![R-CMD-check](https://github.com/jvieroe/epinionDSB/workflows/R-CMD-check/badge.svg)](https://github.com/jvieroe/epinionDSB/actions)
<!-- badges: end -->

# Introduction

The goal of ```epinionDSB``` is to provide easy-to-use templates for data visualizations using the ```ggplot2``` package and in line with the Corporate Visual Identity (CVI) of DSB.

For this purpose, `epinionDSB` functionality includes:

1. a set of tailor-made `ggplot2` themes
2. custom-made color palettes for both discrete and continuous mapping variables


## Quick example
```{r, warning=F, message=F}
library(tidyverse)
library(epinionDSB)

ggplot(mtcars, aes(x = wt,
                   y = mpg)) +
  geom_point(aes(color = factor(cyl)),
             alpha = .85,
             size = 4) +
  facet_wrap(~ vs) +
  dsb_theme_classic(gridlines = "both") +
  color_dsb_d(reverse = T)
  

```


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
if(!require("devtools")) install.packages("devtools")
library(devtools)
devtools::install_github("jvieroe/epinionDSB")
```
## Readymade `ggplot2` themes

`epinionDSB::dsb_theme_*()` adds a tailormade theme to your `ggplot2` objects. This makes them compliant with the DSB CVI and serves as a time-saver when producing several plots.

Choose between three themes:

* `dsb_theme_classic()`, designed for most visualization purposes
  + arguments: `legend`, `gridlines`, `textcolor`
* `dsb_theme_map()`, a very minimalist theme designed with geospatial maps in mind
  + arguments: `legend`, `textcolor`
* `dsb_theme_dust()`, to a large degree similar to `dsb_theme_classic()` but with a warmer, dusty feel
  + arguments: `legend`, `gridlines`, `textcolor`, `background`


The ```dsb_theme_*()``` functions do not impact the `aes()` of your plot, only the `ggplot2::theme()` elements. Note that all `ggplot2::theme()` settings inherent in `dsb_theme*()` can be overwritten by adding `theme(...)` elements after.

```{r cars2}
ggplot(mtcars, aes(x = wt,
                   y = mpg,
                   color = factor(am))) +
  geom_point(size = 4) +
  dsb_theme_classic() +
  theme(legend.position = "right")

```


<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>. -->


<!-- In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->


## DSB color palettes

You can apply DSB themed colors with `epinionDSB`. Choose between the following approaches:

### Extract HEX codes using `dsb_colvec()`
The `epinionDSB` package provides access to the color palette in the DSB design manual. Specifically, the ```dsb_colvec()``` function extracts HEX codes by the color names.

```{r warning=F, message=F}
# ... all DSB colors
dsb_colvec()

# a selection of colors
dsb_colvec("DSB Red", "DSB DarkBlue")
```

We can use these in our `ggplot2` syntax by using the HEX codes or, more elegantly, by using the `base::getElement()` function:

```{r warning=F, message=F}
ggplot(mtcars, aes(x = wt,
                   y = mpg,
                   color = factor(am))) +
  geom_point(size = 4) +
  dsb_theme_classic() +
  scale_colour_manual(values = c(getElement(dsb_colvec(), "DSB Red"), 
                                 getElement(dsb_colvec(), "DSB DarkBlue"))) +
  theme(legend.position = "none")

```

### Apply ready-made color palettes

To provide a more verbose alternative, `epinionDSB` contains out-of-the-box functions to provide our `ggplot2` figures with the DSB color palette. These functions apply to both discrete and continuous variable mapping:


__Discrete variables__

* ```color_dsb_d```: to use with the ```aes(color = x)```, where x is a ```factor``` or ```character``` variable
* ```fill_dsb_d```: to use with the ```aes(fill = x)```, where x is a ```factor``` or ```character``` variable

__Continuous variables__

* ```color_dsb_c```: to use with the ```aes(color = x)```, where x is a ```numeric``` or ```integer``` variable
* ```fill_dsb_c```: to use with the ```aes(fill = x)```, where x is a ```numeric``` or ```integer``` variable


__Let's check it out!__

```{r warning=F, message=F}

# discrete variable in aes()
p1 <-
  ggplot(mtcars, aes(x = wt,
                     y = mpg,
                     color = factor(am))) +
  geom_point(size = 4) +
  dsb_theme_classic(gridlines = "both") +
  theme(legend.position = "right") +
  color_dsb_d()

# continuous variable in aes()
p2 <-
  ggplot(mtcars, aes(x = wt,
                     y = mpg,
                     color = disp)) +
  geom_point(size = 4) +
  dsb_theme_dust() +
  theme(legend.position = "right") +
  color_dsb_c()

library(patchwork)
p1 / p2

```

The main argument taken by all four functions is `reverse` which allows you to reverse the order of the color scale (default is `FALSE`). Furthermore, `*_dsb_c()` allows you to select from a range of color palettes: `reds`, `greens`, `blues`, and `teals`. You choose between these with the `palette` option in `color_dsb_c()` and `fill_dsb_c()`.

```{r warning=F, message=F}
# continuous variable in aes()
p1 <- 
  ggplot(mtcars, aes(x = wt,
                         y = mpg,
                         color = disp)) +
  geom_point(size = 4) +
  dsb_theme_dust(background = "yes") +
  theme(legend.position = "right") +
  color_dsb_c(palette = "greens")

p2 <- 
  ggplot(mtcars, aes(x = wt,
                         y = mpg,
                         color = disp)) +
  geom_point(size = 4) +
  dsb_theme_dust(background = "yes") +
  theme(legend.position = "right") +
  color_dsb_c(palette = "greens",
              reverse = TRUE)

p3 <- 
  ggplot(mtcars, aes(x = wt,
                         y = mpg,
                         color = disp)) +
  geom_point(size = 4) +
  dsb_theme_dust(background = "yes") +
  theme(legend.position = "right") +
  color_dsb_c(palette = "blues")

p4 <- 
  ggplot(mtcars, aes(x = wt,
                         y = mpg,
                         color = disp)) +
  geom_point(size = 4) +
  dsb_theme_dust(background = "yes") +
  theme(legend.position = "right") +
  color_dsb_c(palette = "blues",
              reverse = TRUE)


library(patchwork)
(p1 + p2) /
  (p3 + p4) + 
  plot_annotation(theme = 
                    theme(plot.background = 
                            element_rect(color = NA,
                                         fill = "#E8E1D5",
                                         )
                          )
                  )

```

#### Additional arguments: `*_dsb_d()`

* When mapping `color_dsb_d()` or `fill_dsb_d()` to a variable with __only two levels__, you can manually choose colors with the `primary` and `secondary` arguments


#### Additional arguments

Notice that `*_dsb_c()` and `*_dsb_d()` inherently calls ```ggplot2::scale_*_gradientn()``` and ```ggplot2::discrete_scale()```, respectively. For that reason, additional arguments, such as ```guide```, also apply. See [```ggplot2::scale_colour_gradientn()```](https://tidyverse.github.io/ggplot2-docs/reference/scale_gradient.html "Gradient colour scales") and [```ggplot2::discrete_scale()```](https://ggplot2.tidyverse.org/reference/discrete_scale.html "Discrete scale constructor") for details on additional arguments.



# Acknowledgements

* The R Core Team for developing and maintaining the language
* Hadley Wickham ([hadley](https://github.com/hadley)) and the rest of the team working on the amazing ```ggplot2``` package (and, frankly, the entire ```tidyverse``` ecosystem)
* Garrick Aden-Buie ([gadenbuie](https://github.com/gadenbuie)) and Dr Simon Jackson ([drsimonj](https://github.com/drsimonj)) for inspiration
* Steffen Bank ([steffenbank](https://github.com/steffenbank)) for convincing me to finally try to write my own ```R``` package



