---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%",
  dpi = 300,
  fig.align = "center"
)
#knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

# epinionDSB

<!-- badges: start -->
<!-- badges: end -->

The goal of ```epinionDSB``` is to provide easy-to-use templates for visualizations using the ```ggplot2``` package


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
if(!require("devtools")) install.packages("devtools")
library(devtools)
devtools::install_github("jvieroe/epinionDSB")
```
## Readymade ```ggplot2``` style

The ```epinionDSB::dsb_style()``` function adds a tailormade theme to your ```ggplot2``` graphics. This allows for a coherent Corporate Visual Identity (CVI) for Epinion consultants working with DSB. Importantly, the ```dsb_style()``` function both

* serves as a time-saver when producing many visualizations, as compared to specifying ```ggplot2::theme()``` in each plot
* allows you to efficiently align visual styles and apply these across the board by altering the content of the ```dsb_style()``` function itself

### Example
```{r example, warning=F, message=F}
library(tidyverse)
library(epinionDSB)
```

The ```dsb_style()``` function does not impact the ```aes()``` of your ```geom_*()```, only the ```ggplot2::theme()```. The only argument taken by ```dsb_style``` (at the time of writing) is the ```legend``` option specifying whether a legend is included and defaulting to ```TRUE``` (with ```legend.position``` defaulting to ```"right"```)


```{r cars}
ggplot(mtcars, aes(x = wt,
                   y = mpg,
                   color = factor(am))) +
  geom_point(size = 3) +
  facet_wrap(~ vs) +
  dsb_style()

```



Note that all ```ggplot2::theme()``` settings inherent in ```dsb_style()``` can be overwritten by adding ```theme()``` elements after ```dsb_style()```:

```{r cars2}
ggplot(mtcars, aes(x = wt,
                   y = mpg,
                   color = factor(am))) +
  geom_point(size = 3) +
  facet_wrap(~ vs) +
  dsb_style() +
  theme(legend.position = "bottom")

```


<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>. -->


<!-- In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->


## Color palettes

Above, I get the standard colors in ```ggplot2``` because ```aes(color = factor(am))``` was not followed by a color specification. We could manually add colors as specified by *[the DSB Design Manual](https://github.com/jvieroe/epinionDSB/blob/main/data/DSB_graf-farver.pdf)* (DSBDM). 

### Extract HEX codes using ```dsb_colvec()```
The ```epinionDSB``` package provides access to the color palette in the DSBDM and for readily used color palettes. Specifically, the ```dsb_colvec``` function extracts HEX codes by the color names.

```{r warning=F, message=F}
# ... all DSB colors
dsb_colvec()

# a selection of colors
dsb_colvec("DSB_Red", "DSB_DarkBlue")
```

We can use these in our ```ggplot2``` syntax:

```{r warning=F, message=F}
# ... a selection of colors
ggplot(mtcars, aes(x = wt,
                   y = mpg,
                   color = factor(am))) +
  geom_point(size = 3) +
  facet_wrap(~ vs) +
  dsb_style() +
  scale_colour_manual(values = c("#B41730", "#00233C"))
```

Great! This is **much** more in line with DSB's visual identity. This is a tedious process, however, and any changes to the CVI would involve a lot of manual changes in our accumulated ```ggplot2``` syntax.

### Apply ready-made color palettes

To provide a more verbose alternative, the ```epinionDSB``` contains out-of-the-box syntax to provide our ```ggplot2``` figures with a color palette consisting of the colors in the DSBDM:

* ```color_dsb_d```: to use with the ```aes(color = x)```, where x is a ```factor``` or ```character``` variable
* ```color_dsb_c```: to use with the ```aes(color = x)```, where x is a ```numeric``` or ```integer``` variable
* ```fill_dsb_d```: to use with the ```aes(fill = x)```, where x is a ```factor``` or ```character``` variable
* ```fill_dsb_c```: to use with the ```aes(fill = x)```, where x is a ```numeric``` or ```integer``` variable

Note that, similarly to ```ggplot2::scale_color_*()```, ```color_dsb_*``` is **not** sensitive to American/English spelling and ```colour_dsb_*``` works identically. 

Let's check it out

```{r warning=F, message=F}
library(gridExtra)

# discrete variable in aes()
p1 <-
  ggplot(mtcars, aes(x = wt,
                     y = mpg,
                     color = factor(am))) +
  geom_point(size = 3) +
  facet_wrap(~ vs) +
  dsb_style() +
  color_dsb_d()

# continuous variable in aes()
p2 <-
  ggplot(mtcars, aes(x = wt,
                     y = mpg,
                     color = disp)) +
  geom_point(size = 3) +
  facet_wrap(~ vs) +
  dsb_style() +
  color_dsb_c()

grid.arrange(
  p1, p2,
  ncol = 1
  )
```

```color_dsb_*()``` primarily takes two arguments as inputs: 

* ```palette```, which allows you to choose your color palette. Call ```epinionDSB::dsb_palettes``` for a complete overview 
* ```reverse```, a logical value indicating the scale is reversed (default is ```FALSE```)


**Let's try both out!**

```{r warning=F, message=F}
# continuous variable in aes()
p1 <- 
  ggplot(mtcars, aes(x = wt,
                         y = mpg,
                         color = disp)) +
  geom_point(size = 3) +
  facet_wrap(~ vs) +
  dsb_style() +
  color_dsb_c(palette = "greens")

p2 <- 
  ggplot(mtcars, aes(x = wt,
                         y = mpg,
                         color = disp)) +
  geom_point(size = 3) +
  facet_wrap(~ vs) +
  dsb_style() +
  color_dsb_c(palette = "greens",
              reverse = TRUE)

grid.arrange(
  p1, p2,
  ncol = 1
  )

```


Notice however that ```color_dsb_c()``` and ```color_dsb_d()``` inherently calls ```ggplot2::scale_colour_gradientn()``` and ```ggplot2::discrete_scale()```, respectively. For that reason, additional arguments such as ```guide``` also apply See [```ggplot2::scale_colour_gradientn()```](https://tidyverse.github.io/ggplot2-docs/reference/scale_gradient.html "xxx") and [```ggplot2::discrete_scale()```](https://ggplot2.tidyverse.org/reference/discrete_scale.html "xxx") for details on additional arguments.


