---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%",
  dpi = 300,
  fig.align = "center"
)
#knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

# epinionDSB

<!-- badges: start -->
<!-- badges: end -->

The goal of ```epinionDSB``` is to provide easy-to-use templates for visualizations using the ```ggplot2``` package


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jvieroe/epinionDSB")
```
## Readymade ```ggplot2``` style

The ```epinionDSB::dsb_style()``` function adds a tailormade theme to your ```ggplot2``` graphics. This allows for a coherent Corporate Visual Identity (CVI) for Epinion consultants working with DSB. Importantly, the ```dsb_style()``` function both

* serves as a time-saver when producing many visualizations, as compared to specifying ```ggplot2::theme()``` in each plot
* allows you to efficiently align visual styles and apply these across the board by altering the content of the ```dsb_style()``` function itself

### Example: ```dsb_style()```
```{r example, warning=F, message=F}
library(tidyverse)
library(epinionDSB)
```


```{r cars}
ggplot(mtcars, aes(x = wt,
                   y = mpg,
                   color = factor(am))) +
  geom_point(size = 3) +
  facet_wrap(~ cyl) +
  dsb_style()

```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
